---
title: "p8105_hw2_yq2279"
author: "Qi Yuchen"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

### Read and clean the Mr. Trash Wheel sheet

```{r}
data_trash = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N336") %>% 
  drop_na(Dumpster) %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))

```

### Read and clean precipitation data for 2017 and 2018

```{r}
# read precipitation data for 2017
data_pre2017 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>% 
  drop_na(Total) %>%   # omit rows without precipitation data
  janitor::clean_names() %>% 
  mutate(year = 2017)   # add a variable year

# read precipitation data for 2018
data_pre2018 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, range = "A2:B14") %>% 
  drop_na(Total) %>%   # omit rows without precipitation data
  janitor::clean_names() %>% 
  mutate(year = 2018)   # add a variable year

# combine precipitation datasets and convert month to a character variable
data_pre2017_2018 = bind_rows(data_pre2017, data_pre2018) %>% 
  mutate(month = month.name[month])

```

* In the data from the Mr. Trash Wheel sheet, the number of observations is `r nrow(data_trash)`, and the key variables include weight_tons, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, sports_balls, etc. 
* In the precipitation data for 2017 and 2018, the number of observations is `r nrow(data_pre2017_2018)`, and the key variables include month and total. 
* The total precipitation in 2018 is `r filter(data_pre2017_2018, year == 2018) %>% pull(total) %>% sum()`.
* The median number of sports balls in a dumpster in 2017 is `r filter(data_trash, year == 2017) %>% pull(sports_balls) %>% median()`.


# Problem 2

### clean the data in pols-month.csv

```{r}
data_pols_month = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  pivot_longer(cols = starts_with("prez"),
               names_to = "president",
               names_prefix = "prez_") %>%
  filter(value != 0) %>% 
  select(-value) %>% 
  select(-day)
```

### Clean the data in snp.csv

```{r}
data_snp = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(year = as.numeric(year), month = as.numeric(month)) %>% 
  arrange(year, month) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, everything()) %>% 
  select(-day) %>% 
  mutate(year = as.character(year))
```

### Tidy the unemployment data 
```{r}
data_unemployment = read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(cols = jan:dec,
               names_to = "month",
               values_to = "unemployment") %>%
  arrange(year, month) %>%
  mutate(month = month.name[factor(month)]) %>% 
  mutate(year = as.character(year))
```

### Join the datasets by merging snp into pols, and merging unemployment into the result
```{r}
pols_snp_unemployment = left_join(data_pols_month, data_snp, by = c("year", "month")) %>% 
  left_join(data_unemployment, by = c("year", "month"))
```

* In data_pols_month, the number of observations is `r nrow(data_pols_month)` and there are `r ncol(data_pols_month)` variables. The range of year is from `r range(pull(data_pols_month, year))[1]` to `r range(pull(data_pols_month, year))[2]`. The key variables are `r names(data_pols_month[3:9])`.
* In data_snp, the number of observations is `r nrow(data_snp)` and there are `r ncol(data_snp)` variables. The range of year is from `r range(pull(data_snp, year))[1]` to `r range(pull(data_snp, year))[2]`. The key variable is `r names(data_snp[3])`.
* In data_unemployment, the number of observations is `r nrow(data_unemployment)` and there are `r ncol(data_unemployment)` variables. The range of year is from `r range(pull(data_unemployment, year))[1]` to `r range(pull(data_unemployment, year))[2]`. The key variable is `r names(data_unemployment[3])`.
* In the joint dataset pols_snp_unemployment, the number of observations is `r nrow(pols_snp_unemployment)` and there are `r ncol(pols_snp_unemployment)` variables. The range of year is from `r range(pull(pols_snp_unemployment, year))[1]` to `r range(pull(pols_snp_unemployment, year))[2]`. The key variables are `r names(pols_snp_unemployment[3:11])`.


# Question 3

### Load and tidy the data on the popularity of baby names
```{r}
data_baby_names = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    ethnicity = recode(ethnicity, 
                       "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                       "BLACK NON HISP" = "BLACK NON HISPANIC",
                       "WHITE NON HISP" = "WHITE NON HISPANIC"),
    ethnicity = str_to_lower(ethnicity)
    ) %>% 
  mutate(
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_lower(gender)
    ) %>% 
  distinct()
```

### Produce a table showing the rank of “Olivia”
```{r}
data_Olivia = 
  filter(data_baby_names, gender == "female" & childs_first_name == "Olivia") %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank",
    id_cols = ethnicity
    )

data_Olivia
```

### Produce a table showing the most popular male name
```{r}
data_male = 
  filter(data_baby_names, gender == "male" & rank == 1) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name",
    id_cols = ethnicity
    )

data_male
```

### Produce a scatter plot about male white non-hispanic children born in 2016
```{r}
data_plot = filter(data_baby_names, 
                   gender == "male" & 
                   ethnicity == "white non hispanic" & 
                   year_of_birth == 2016)
ggplot(data_plot, aes(x = rank, y = count)) +
  geom_point() +
  xlab("Rank in Popularity") +
  ylab("Number of Children") 
```

